\bf Title here

### 2020-08-20 paper/paper.Rmd
motiational paragraph for data structure; mention of tidy data at the end of Sec 4.1

# Visual diagnostics{#vis-diag}

@@

There are different ways to represent the same data in rectangular form. Certain cleaning steps are needed, before using ggplot2, to bring the data into a tidy data format [@wickham2014tidy] where 1) each observation forms a row; 2) each variable forms a column and 3) each type of observational unit forms a table. In a tidy format, data wrangling and visualisation are greatly simplified.

@@

Global object is a data construction with key components from optimisation algorithms and is designed for making diagnostic plots easier. In the optimisation algorithms for projection pursuit, three key elements are 1) projection bases: $\mathbf{A}$, 2) index values: $I$ and 3) State: $S$, which labels the observation with detailed stage of searching or interpolation.

Multiple iterators are needed to index the data collected at different levels. $t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts. $l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output.

@@

There are other parameters that are of our interest and we denote them as *$V_{p}$*. In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases. The data structure can thus be shown as in Equation \ref{eq:data-structure}.

@@

A sketch of the global object for projection pursuit guided tour is presented in Figure \ref{fig:glb-obj}.

[I feel this sketch was initially useful but now since we have the data matrix and the printed output of the global object, it doesn't any additional information. I'm still keeping it here but we may need to remove it if it's not useful :(]

```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The global object in projection pursuit guided tour."}
knitr::include_graphics(path = here::here("figures/global_obj.png"))
```

@@

<!-- Since the global object is already tidy, not further tidying steps is needed, while certain data wrangling steps [@wickham2016rfordatascience] are still needed to transform the global object into  desirable format for one particular visualisation. To emphasize on this good practice of data analysis, we will describe the transformation steps needed for each diagnostic plots before stepping into visualisation. -->

@@

We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the global object.

@@

- Layer 1: point geom
  - data: filter the observations with  $S = \text{interpolation}$; bind the global object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
  - x: $t$ is mapped to the x-axis
  - y: $I$ is mapped to the y-axis 
  - colour: $V$ is mapped to the colour aesthetic

@@

When the optimisation ends, the global object will be stored and printed (it can be turned off by supplying argument `print = FALSE`). Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.  Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.

>>>>

# Visual diagnostics {#vis-diag}

To be able to make diagnostic plot, the optimisation algorithm should populate a data structure that contains the key elements of the algorithm. When the algorithm runs, key information regarding the decision variable, objective function and hyper-parameters needs to be recorded and stored as a data object for future analysis.

@@

@@

In the optimisation algorithms for projection pursuit, three main elements to record are 1) projection bases: $\mathbf{A}$, 2) index values: $I$, and 3) State: $S$, which labels the observation with detailed stage in the optimisation. Multiple iterators are also  needed to index the data collected at different levels. $t$ is a unique identifier that prescribes the natural ordering of each observation; $j$ is the counter for each search-and-interpolate round, which remains the same within one round and has an increment of one once a new round starts. $l$ is the counter for each search/interpolation allowing us to know how many basis the algorithm has searched before finding one to output. There are other parameters that are of our interest and we denote them as *$V_{p}$*. In projection pursuit, this includes $V_1 = \text{method}$, which tags the name of the algorithm used and $V_2 = \text{alpha}$, the neighbourhood parameter that controls the size in sampling candidate bases.  A matrix notation of the data structure is presented in Equation \ref{eq:data-structure}.

@@

@@

```{r glb-obj, out.width= "100%", out.height="20%", fig.cap="\\label{glb-obj}The data structure in projection pursuit guided tour."}
knitr::include_graphics(path = here::here("figures/global_obj.png"))
```

It is worth-noticing that the data structure constructed above meets the tidy data principal [@wickham2014tidy] that states

1) each observation forms a row, 
2) each variable forms a column, and 
3) each type of observational unit forms a table

The wrangling and visualisation of tidy data have been greatly simplified by the well-known dplyr[@dplyr] and ggplot2[@ggplot2] package.

@@

@@

We form our first dataset using variables `x1`, `x2`, `x8`, `x9` and `x10`,  run the guided tour with optimiser `search_better` and below shows the first ten rows of the data object.

@@

- Layer 1: point geom
  - data: filter the observations with  $S = \text{interpolation}$; bind the data object from optimisation and interpolation and form polishing; mutate $t$ to be the row number of the binded tibble.
  - x: $t$ is mapped to the x-axis
  - y: $I$ is mapped to the y-axis 
  - colour: $V$ is mapped to the colour aesthetic

@@

When the optimisation ends, the data object will be stored and printed (it can be turned off by supplying argument `print = FALSE`). Additional messages during the optimisation can be displayed by argument `verbose = TRUE`.  Notice that the tibble object allows the list-column `basis` to be printed out nicely with the dimension of the projection basis readily available.

### 2020-08-20 paper/paper.tex
motiational paragraph for data structure; mention of tidy data at the end of Sec 4.1

@@

There are different ways to represent the same data in rectangular form.
Certain cleaning steps are needed, before using ggplot2, to bring the
data into a tidy data format \citep{wickham2014tidy} where 1) each
observation forms a row; 2) each variable forms a column and 3) each
type of observational unit forms a table. In a tidy format, data
wrangling and visualisation are greatly simplified.

Global object is a data construction with key components from
optimisation algorithms and is designed for making diagnostic plots
easier. In the optimisation algorithms for projection pursuit, three key
elements are 1) projection bases: \(\mathbf{A}\), 2) index values: \(I\)
and 3) State: \(S\), which labels the observation with detailed stage of
searching or interpolation.

Multiple iterators are needed to index the data collected at different
levels. \(t\) is a unique identifier that prescribes the natural
ordering of each observation; \(j\) is the counter for each
search-and-interpolate round, which remains the same within one round
and has an increment of one once a new round starts. \(l\) is the
counter for each search/interpolation allowing us to know how many basis
the algorithm has searched before finding one to output.

There are other parameters that are of our interest and we denote them
as \emph{\(V_{p}\)}. In projection pursuit, this includes
\(V_1 = \text{method}\), which tags the name of the algorithm used and
\(V_2 = \text{alpha}\), the neighbourhood parameter that controls the
size in sampling candidate bases. The data structure can thus be shown
as in Equation \ref{eq:data-structure}.

@@

A sketch of the global object for projection pursuit guided tour is
presented in Figure \ref{fig:glb-obj}.

{[}I feel this sketch was initially useful but now since we have the
data matrix and the printed output of the global object, it doesn't any
additional information. I'm still keeping it here but we may need to
remove it if it's not useful :({]}

\begin{figure}
\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The global object in projection pursuit guided tour.}\label{fig:glb-obj}
\end{figure}

@@

We form our first dataset using variables \texttt{x1}, \texttt{x2},
\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
optimiser \texttt{search\_better} and below shows the first ten rows of
the global object.

@@

  \begin{itemize}
  \tightlist
  \item
    data: filter the observations with \(S = \text{interpolation}\);
    bind the global object from optimisation and interpolation and form
    polishing; mutate \(t\) to be the row number of the binded tibble.
  \item
    x: \(t\) is mapped to the x-axis
  \item
    y: \(I\) is mapped to the y-axis
  \item
    colour: \(V\) is mapped to the colour aesthetic
  \end{itemize}
\item
  Layer 2: line geom
\end{itemize}

@@

When the optimisation ends, the global object will be stored and printed
(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
Additional messages during the optimisation can be displayed by argument
\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
list-column \texttt{basis} to be printed out nicely with the dimension
of the projection basis readily available.

>>>>

To be able to make diagnostic plot, the optimisation algorithm should
populate a data structure that contains the key elements of the
algorithm. When the algorithm runs, key information regarding the
decision variable, objective function and hyper-parameters needs to be
recorded and stored as a data object for future analysis.

@@

In the optimisation algorithms for projection pursuit, three main
elements to record are 1) projection bases: \(\mathbf{A}\), 2) index
values: \(I\), and 3) State: \(S\), which labels the observation with
detailed stage in the optimisation. Multiple iterators are also needed
to index the data collected at different levels. \(t\) is a unique
identifier that prescribes the natural ordering of each observation;
\(j\) is the counter for each search-and-interpolate round, which
remains the same within one round and has an increment of one once a new
round starts. \(l\) is the counter for each search/interpolation
allowing us to know how many basis the algorithm has searched before
finding one to output. There are other parameters that are of our
interest and we denote them as \emph{\(V_{p}\)}. In projection pursuit,
this includes \(V_1 = \text{method}\), which tags the name of the
algorithm used and \(V_2 = \text{alpha}\), the neighbourhood parameter
that controls the size in sampling candidate bases. A matrix notation of
the data structure is presented in Equation \ref{eq:data-structure}.

@@

\begin{figure}
\includegraphics[width=1\linewidth,height=0.2\textheight]{/Users/hzha400/Documents/3.PhD/research/paper-tour-vis/figures/global_obj} \caption{\label{glb-obj}The data structure in projection pursuit guided tour.}\label{fig:glb-obj}
\end{figure}

It is worth-noticing that the data structure constructed above meets the
tidy data principal \citep{wickham2014tidy} that states

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  each observation forms a row,
\item
  each variable forms a column, and
\item
  each type of observational unit forms a table
\end{enumerate}

The wrangling and visualisation of tidy data have been greatly
simplified by the well-known dplyr\citep{dplyr} and
ggplot2\citep{ggplot2} package.

@@

We form our first dataset using variables \texttt{x1}, \texttt{x2},
\texttt{x8}, \texttt{x9} and \texttt{x10}, run the guided tour with
optimiser \texttt{search\_better} and below shows the first ten rows of
the data object.

@@

  \begin{itemize}
  \tightlist
  \item
    data: filter the observations with \(S = \text{interpolation}\);
    bind the data object from optimisation and interpolation and form
    polishing; mutate \(t\) to be the row number of the binded tibble.
  \item
    x: \(t\) is mapped to the x-axis
  \item
    y: \(I\) is mapped to the y-axis
  \item
    colour: \(V\) is mapped to the colour aesthetic
  \end{itemize}
\item
  Layer 2: line geom
\end{itemize}

@@

When the optimisation ends, the data object will be stored and printed
(it can be turned off by supplying argument \texttt{print\ =\ FALSE}).
Additional messages during the optimisation can be displayed by argument
\texttt{verbose\ =\ TRUE}. Notice that the tibble object allows the
list-column \texttt{basis} to be printed out nicely with the dimension
of the projection basis readily available.
